---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Names of your team members

### Date

Winter 2022

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Include a chart

Describe why you chose this chart and what information it reveals

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart1_example.R")}
```

### Chart 2

```{r}

```


### Chart 3

```{r}
source("./anaylsis.R")

Phoniex_plot
```

```{r}
source("./anaylsis.R")

Astoria_plot
```

# Chart 3- Amen

For these graphs, we chose to focus on cities with the highest and lowest temperature changes. We specifically targeted the western region, selecting Arizona and Oregon—Arizona being in the Southwest and Oregon in the West. The average temperature of Phoenix was included because it shows significant temperature changes since the early 1900s, making it one of the hottest cities. In contrast, the average temperature of Astoria, which has the lowest temperature change, was also included. We wanted to compare these two cities, given their relatively close geographic proximity within the region, to highlight the stark differences in their temperature trends.


# Includes insights (what information is revealed)


In the chart showing the average temperature of Phoenix, we observe that the temperature hasn't dipped below 65 degrees for over 120 years, consistently maintaining a warm climate. Over time, this temperature has continued to rise.


In contrast, the chart displaying the average temperature of Astoria reveals significantly lower temperatures compared to Phoenix. The average temperature in Astoria ranges from 45 to 55 degrees. Although there is a dip at some point, the temperature has generally remained within this range.


# Visualization type makes sense for data


Line charts makes sense for the data because they best represent the information we want to share clearly, especially since we are comparing data with high and low values.
